---
title: "Data Visualiation with R Exercises"
mainfont: Calibri Light
sansfont: Calibri
monofont: Courier New
output:
  word_document: default
  pdf_document:
    latex_engine: xelatex
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r load_libraries, include=FALSE, message=FALSE}
# load packages
packages <-  c("knitr", "rprojroot")
for( package in packages ){
  library( package, character.only = TRUE )
}
```

```{r knitr_options, include=FALSE, message=FALSE}
opts_chunk$set( fig.width=9, fig.height=6, include=FALSE, echo=FALSE, 
                results = FALSE, warning = FALSE, message = FALSE,
                fig.show = FALSE, cache = FALSE)

```

Load the example dataset (`example2.tsv`) using `read_tsv`.

```{r tidyverse, include = TRUE, echo = TRUE, eval = FALSE}
library(tidyverse)
results <- read_tsv('example2.tsv',
                    col_types = cols(Chr = 'c', Strand = 'c'))
```

```{r load_data, include = TRUE, echo = FALSE}
library(tidyverse)
results <- read_tsv('r-data-vis/example2.tsv',
                    col_types = cols(Chr = 'c', Strand = 'c'))
```

## Filtering and Selecting

1. Use `filter` to find out how many genes have an adjusted p-value less than 0.05.

```{r filtering}
filter(results, adjp < 0.05) %>%
  nrow()
```

2. Find out which gene has the smallest p-value by sorting the data using `arrange`.

```{r arrange}
arrange(results, adjp) %>%
  select(., GeneID, Name, adjp) %>%
  head(1)
```

3. Make a new column in the data that is -log10 of the adjusted p-value column.
You can use the `log10()` function to calculate this.

```{r mutate}
results <- mutate(results, 
       log10p = -log10(adjp))

select(results, GeneID, Name, adjp, log10p) %>% 
  arrange(adjp) %>% 
  head(5)
```

4. Make a new data.frame that contains the GeneID, Name and all the normalised
count columns.

```{r select}
normalised_counts <- 
  select(results, GeneID, Name, contains('normalised_count'))
```

5. Make the new data.frame data tidy using `pivot_longer`.

```{r pivot, eval = FALSE}
normalised_counts_long <- 
  pivot_longer(normalised_counts, c(-GeneID, -Name),
                names_to = 'sample', values_to = 'normalised count')
```

```{r gather, echo = FALSE}
normalised_counts_long <- 
  gather(normalised_counts, c(-GeneID, -Name),
          key = 'sample', value = 'normalised count')

```

## Plotting

1. Using the `iris` dataset make a plot of Petal.Width (y) against Sepal.Width (x).

```{r plot1}
ggplot(data = iris) + 
  geom_point(aes(x = Sepal.Width, y = Petal.Width))
```

2. Now colour the points by Petal.Length and use the viridis colour scale 
using `scale_colour_viridis_c`.

```{r plot2, eval = FALSE}
ggplot(data = iris, aes(x = Sepal.Width, y = Petal.Width)) + 
  geom_point(aes(colour = Petal.Length)) +
  scale_colour_viridis_c()
```

```{r plot2_actual, echo = FALSE}
library(viridis)
ggplot(data = iris, aes(x = Sepal.Width, y = Petal.Width)) + 
  geom_point(aes(colour = Petal.Length)) +
  scale_colour_viridis()
```

3. Change the colouring to Species, choose a hollow shape (one of 21-25) and 
pick 3 colours to use with `scale_fill_manual`.

```{r plot3}
ggplot(data = iris, aes(x = Sepal.Width, y = Petal.Width)) + 
  geom_point(aes(fill = Species), shape = 22, size = 3) +
  scale_fill_manual(values = c('firebrick2', 'steelblue3', 'orange'))
```
